\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-10 casestudies.tex
More Comments.

In addition to the two case studies above, we also use extensible
inductive types for modeling extensible context-free grammars and derive
decision procedures for them.

>>>>

In addition to the two case studies above, we also use extensible
inductive types for modeling extensible context-free grammars and derive
decision procedures for them.\EDJ{I will suggest change to ``and derive decision procedures for them (based on naive parsing by enumeration)''.}

### 2022-11-10 cc-condensed.tex
More Comments.

\noindent
Consistency says that the type $\bot$ is not inhabited.
Thus, it is safe to use the type theory for logical reasoning, as not
every proposition is trivially provable.
\cref{thm:consistency} is a consequence of a meta\-circular interpretation of all
the well-formedness rules, which, in particular, interprets $\bot$ in
\TT as an uninhabited type in the meta\-language.
The construction of this interpretation follows the standard model of
\citet{altkap2016}, extending it to handle linkages.
Linkages~$\LkgAdd{\lkg}{t}$ can be interpreted into non-dependent products,
as rule \ruleref{l/add} indicates.
%
We refer the reader to a supplemental file (in Agda syntax) for the complete proof.

@@

\noindent
This canonicity theorem says that every closed term of the ground
type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
When canonicity is proved in a constructive meta\-logic, its proof
amounts to a normalization function for closed terms of the ground type.\EDJ{Suggesting use "boolean type". Because natural number type can be also considered ground type in other context (but this theorem on boolean cannot ensure canonicity for nat). We also does specify what is ground type. }\YZreply{I think it is clear from the context and from the article 'the' that 'the ground type' refers to B.}
So canonicity serves to justify the computational nature of the type theory.

We prove \cref{thm:canonicity} by constructing a logical-relations model
for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,sterling2019algebraic,kaposi2019gluing}.
In particular, a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
into a logical predicate on closed terms: the predicate includes all
closed, ``reducible'' terms of type~$T$.\EDJ{I will add a it is my bad to introduce "reducible" without defining it first. But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)}\YZreply{Is there a more accurate way to say things here?}
The proof is available as a supplemental file in Agda syntax.

>>>>

\noindent
Consistency says that the type $\bot$ is not inhabited.
Thus, it is safe to use the type theory for logical reasoning, as not
every proposition is trivially provable.
\cref{thm:consistency} is a consequence of a meta\-circular interpretation of all
the well-formedness rules, which, in particular, interprets $\bot$ in
\TT as an uninhabited type in the meta\-language.
The construction of this interpretation follows the standard model of
\citet{altkap2016}, extending it to handle linkages.
Linkages~$\LkgAdd{\lkg}{t}$ can be interpreted into non-dependent products\EDJ{Suggesting changing to ``can be interpreted into non-dependent pair with dependent function for late-binding''},
as rule \ruleref{l/add} indicates.
%
We refer the reader to a supplemental file (in Agda syntax) for the complete proof.

@@

\noindent
This canonicity theorem says that every closed term of the ground
type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
When the canonicity theorem is proved in a constructive meta\-logic, its proof
amounts to a normalization function for closed terms of the ground type.\EDJ{Suggesting use "boolean type". Because natural number type can be also considered ground type in other context (but this theorem on boolean cannot ensure canonicity for nat). We also does specify what is ground type. }\YZreply{I think it is clear from the context and from the article 'the' that 'the ground type' refers to B.}
So canonicity serves to justify the computational nature of the type theory.

We prove \cref{thm:canonicity} by constructing a logical-relations model
for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,sterling2019algebraic,kaposi2019gluing}.
In particular, a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
into a logical predicate on closed terms: the predicate includes all
closed, ``reducible'' terms of type~$T$.\EDJ{I will add a quote around reducible here. it is my bad to introduce "reducible" without defining it first. But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)}\YZreply{Is there a more accurate way to say things here?}\EDJreply{Because in the operational semantic, using ``reducible'' is totally fine. But in declarative style, only ``canonicity'' this word is making sense. But this ``canonicity'' is not friendly to the audience not familar with declartive style.}
The proof is available as a supplemental file in Agda syntax.

### 2022-11-10 coqimpl.tex
Some Comments.
More Comments.

Importantly,
\lsti{STLC\_tm} merely declares the existence of these names and their types;
it does not specify their definitions or proofs.
Knowing the existence of these names and their types
suffices to type-check and translate the rest of the family.
Leaving the definitions and proofs open enables \lsti{STLC} and \lsti{STLCFix}
to instantiate \lsti{tm} differently upon \lsti{End}~\lsti{STLC} and \lsti{End}~\lsti{STLCFix}.

@@

Upon \lsti{End}~\lsti{subst}, a module type named \lsti{STLC\_subst} is generated.
%
As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
so its definition is not exposed to the fields that come after it.
Accordingly, the translation \lsti{STLC\_subst} merely declares the type of \lsti{subst}
and the existence of its propositional equalities, leaving \lsti{subst} undefined
and the equalities unproven.
%
The propositional equalities are stated in terms of the case handlers,
whose definitions \emph{are} available through the \lsti{self} parameter.
So Coq can simplify, for example, \lsti{subst_tm_unit_eq}'s type to
\lsti!\forall x t, self.subst self.tm_unit x t = self.tm_unit!.
The equalities will be included and available for use in the
translations of the subsequent fields through their \lsti{self}
parameters.

>>>>

Importantly,
\lsti{STLC\_tm} merely declares the existence of these names and their types;\EDJ{What does this sentence mean?}
it does not specify their definitions or proofs.
Knowing the existence of these names and their types
suffices to type-check and translate the rest of the family.
Leaving the definitions and proofs open enables \lsti{STLC} and \lsti{STLCFix}
to instantiate \lsti{tm} differently upon \lsti{End}~\lsti{STLC} and \lsti{End}~\lsti{STLCFix}.

@@

Upon \lsti{End}~\lsti{subst}, a module type named \lsti{STLC\_subst} is generated.
%
As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,\EDJ{If we don't allow further bound of case handler(mentioned in your last reply), how do the user further bound the subst? I mean in the surface syntax}
so its definition is not exposed to the fields that come after it.
Accordingly, the translation \lsti{STLC\_subst} merely declares the type of \lsti{subst}
and the existence of its propositional equalities, leaving \lsti{subst} undefined
and the equalities unproven.
%
The propositional equalities are stated in terms of the case handlers,
whose definitions \emph{are} available through the \lsti{self} parameter.
So Coq can simplify, for example, \lsti{subst_tm_unit_eq}'s type to
\lsti!\forall x t, self.subst self.tm_unit x t = self.tm_unit!.
The equalities will be included and available for use in the
translations of the subsequent fields through their \lsti{self}
parameters.

### 2022-11-10 metatheory-condensed.tex
More Comments.

\TT is intended as a foundational model. So unlike our programmer-facing plugin,
typing is structural rather than nominal, and
fields automatically axiomatized by the plugin require explicit definitions in \TT.
%
\TT provides MLTT-style constructs that can be used to express families and
family polymorphism. %rather than building it in directly.
%
Most notably, it extends MLTT with what we call \emph{linkages}.
(Linkages are a namesake of the theoretical device through which prior work~\cite{zm2017} models
family polymorphism in an OO setting, but the technical details differ significantly from the prior work.)

@@

In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, A}{t}{T}$ is
responsible for late binding.
Here, $A$ abstracts the fields preceding the current field~$t$
(i.e., those recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$).
Crucially, unlike the second field in a dependent pair (\ruleref{tm/pair}),
the typing of~$t$ in \ruleref{l/add} is polymorphic to how the fields in
$t$'s context are defined, so that a different linkage that overrides
fields in $t$'s context can reuse~$t$---without retyping it---by
projecting it and then including it.

This type~$A$ abstracting the context in \ruleref{l/add} and
\ruleref{lsig/add} is not necessarily the same as those recorded by~$\lsig$,
because a field representing the code $\wcode{\wsig}$ of a W-type has to
expose different types to different fields that come after it.
Later fields that invoke
$\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
and \ruleref{tm/wrec} stipulate, whereas $\wsig$ should be hidden
from all other fields so that they can be reused
for a W-type signature~$\wsig'$ extending~$\wsig$.

>>>>

\TT is intended as a foundational model. So unlike our programmer-facing plugin,
typing is structural rather than nominal\EDJ{I don't actually know what do you mean by structural here. I think surface syntax (Coq) is pretty structual as well.}, and
fields automatically axiomatized by the plugin require explicit definitions in \TT.
%
\TT provides MLTT-style constructs that can be used to express families and
family polymorphism. %rather than building it in directly.
%
Most notably, it extends MLTT with what we call \emph{linkages}.
(Linkages are a namesake of the theoretical device through which prior work~\cite{zm2017} models
family polymorphism in an OO setting, but the technical details differ significantly from the prior work.)

@@

In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, A}{t}{T}$ is
responsible for late binding.
Here, $A$ abstracts the fields preceding the current field~$t$
(i.e., those recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$).
Crucially, unlike the second field in a dependent pair (\ruleref{tm/pair}),
the typing of~$t$ in \ruleref{l/add} is polymorphic to how the fields in
$t$'s context are defined, so that a different linkage that overrides
fields in $t$'s context can reuse~$t$---without retyping it---by
projecting it and then including it\EDJ{I suggest change this sentence to `by projecting it out of parent linkage and adding it in the derived linkage'}.

This type~$A$ abstracting the context in \ruleref{l/add} and
\ruleref{lsig/add} is not necessarily the same as those recorded by~$\lsig$\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
because a field representing the code $\wcode{\wsig}$ of a W-type has to
expose different types to different fields that come after it.
Later fields that invoke
$\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
and \ruleref{tm/wrec} stipulate, whereas $\wsig$ should be hidden
from all other fields so that they can be reused
for a W-type signature~$\wsig'$ extending~$\wsig$\EDJ{Suggest changing to ``so that they can be reused
in a different context where W-type signature~$\wsig'$ is extended to~$\wsig$''}.
